---
title: "'dockr': easy containerization for R"
output:
  html_document: default
  pdf_document: default
---
```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# dockr

`dockr` 0.8.5 is now available on CRAN. `dockr` is a minimal toolkit to build a 
lightweight Docker container image for your R package, in which the package 
itself is available. The Docker image seeks to mirror your R session as close as 
possible with respect to R specific dependencies. Both dependencies on CRAN 
R packages as well as local non-CRAN R packages will be included in the Docker 
container image.

If you want to know, how Docker works, and why you should consider using Docker,
please take a look at the [Docker website](https://www.docker.com/why-docker).

## Installation

Install the development version of `dockr` with:

```{r install_github, eval = FALSE}
remotes::install_github("smaakage85/dockr")
```

Or install the version released on CRAN: 

```{r install_cran, eval = FALSE}
install.packages("dockr")
```

## Workflow

When you work on an R project, it is often desirable to organize the code in the
R package structure. `dockr` facilitates easy creation of a Docker container 
image that mirrors your current R session and includes all of the R dependencies 
needed to run your R package. 

First, load the `dockr` package.

```{r load_package, eval = TRUE}
library(dockr)
```

```{r setup_cran, echo = FALSE}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```

In order do create the files, that constitute the Docker image, simply invoke 
the `prepare_docker_image()` function and point to the folder with your package. 

The workflow of `prepare_docker_image()` is summarized below:

1. Build and install the package on your system
2. Identify R package dependencies of the package
3. Detect the version numbers of the loaded and installed versions of these 
packages on your system
4. Write Dockerfile and create all other files needed to build the Docker image

Now, I will let `dockr` do its magic and create the files for a Docker image
container, in which `dockr` is installed together with all of the R package 
dependencies, `dockr` needs to run.

Beware that the files are created as side-effects of the function call.

```{r create_image, warning = FALSE, message = FALSE}
image_dockr <- prepare_docker_image(".")
```

Great, all necessary files for the Docker image have been created, and you 
can build the Docker image right away by following the instructions. It is as 
easy as that! Yeah!

## Files for Docker image

Let us just take a quick look into the folder with the files for the Docker 
image to see the works of `dockr`.

```{r lookup_files_image}
list.files(image_dockr$paths$dir_image)
```

It contains a Dockerfile and a folder named 'source_packages'.

### Dockerfile

The resulting Dockerfile can be printed with the `print_file()` function, that
comes with `dockr`:

```{r print_dockerfile}
print_file(image_dockr$paths$path_Dockerfile)
```

As you see, the versions of the R packages, that will be installed in the Docker 
container image, are all given explicitly. They will mirror the versions of the
dependencies, that are in fact loaded or installed on your system. In this way,
the Docker container image seeks to reflect your current R session as close as 
possible and by doing so create an environment, where you will be able to 
reproduce results from your current R session. 

Also note, that CRAN R packages will be installed from relevant 
[MRAN snapshots](https://mran.microsoft.com/) - using the 
`remotes::install_version()` function.

### Folder with Source Packages

The 'source_packages' folder contains the local (non-CRAN) packages, that have 
to be installed in the Docker container image in order for `dockr` to run.

Since `dockr` does not depend on any local (non-CRAN) packages, 
`source_packages` only contains a source package version of `dockr` itself, 
i.e.:

```{r lookup_files_src}
list.files(image_dockr$paths$dir_source_packages)
```

## How to edit Dockerfile further

If there is need for adding additional lines to/editing the Dockerfile (e.g.
if you have to install any non-R [dependencies](#nonr), this can 
be achieved with the `write_lines_to_file()` function. `write_lines_to_file()` 
enables you to add new lines to the beginning or the end of the Dockerfile. 

Let us try it out and write a couple of additional lines to the Dockerfile.

```{r add_lines}
# write three lines to beginning of file.
write_lines_to_file(c("# set maintainer",
                    "MAINTAINER Lars KJELDGAARD <lars_kjeldgaard@hotmail.com>", 
                    ""),
                    image_dockr$paths$path_Dockerfile,
                    prepend = TRUE,
                    print_file = FALSE)

# write lines to the end of the file.
write_lines_to_file(c("# check out smaakage85.netlify.com >:-]~~"),
                    image_dockr$paths$path_Dockerfile,
                    prepend = FALSE,
                    print_file = FALSE)
```

Take a look at the resulting Dockerfile.

```{r print_dockerfile_again}
print_file(image_dockr$paths$path_Dockerfile)
```

## Dealing with local non-CRAN R package dependencies

If your package depends on local non-CRAN R packages, `dockr` will also include
these packages in the Docker container image. Local non-CRAN R packages must be
available as source packages (*[packageName]_[packageVersion].tar.gz*) in one
or more user specified local directories. These paths have to be specified in the
'dir_src' argument, when invoking the `prepare_docker_image()`, e.g.: 

```{r image_recorder, eval = FALSE}
# image for my package 'recorder'.
image_recorder <- prepare_docker_image("~/recorder",
                                       dir_src = "~/src")
```

## What about non-R dependencies?

`dockr` does _not_ deal with any non-R dependencies what so ever at this point. 
In case that, for instance, your package has any Linux specific dependencies, 
you will have to install them yourself in the Docker container image. 

## Contact

I hope, that you will find `dockr` useful.

Please direct any questions and feedbacks to [me](mailto:lars_kjeldgaard@hotmail.com)!

If you want to contribute, open a [PR](https://github.com/smaakage85/dockr/pulls).

If you encounter a bug or want to suggest an enhancement, please [open an issue](https://github.com/smaakage85/dockr/issues).

Best, 
smaakagen









