---
title: "dockr hits CRAN"
output:
  html_document: default
  pdf_document: default
---
---
title: "'dockr' hits CRAN"

---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# dockr

`dockr` 0.8.0 is now available on CRAN. `dockr` is a minimal toolkit to build a 
lightweight Docker image from your R package, in which the package itself is 
installed and available. The image seeks to mirror your R session as close as 
possible with respect to R specific dependencies.

## Installation

Install the development version of `dockr` with:

```{r, eval = FALSE}
remotes::install_github("smaakage85/dockr")
```

Or install the version released on CRAN: 

```{r, eval = FALSE}
install.packages("dockr")
```


```{r setup, eval = TRUE}
library(dockr)
```

## Workflow

When you work on an R project, it is often desirable to organize the code in an
R package. `dockr` prepares a Docker container image that mirrors your R session
and includes all of the R dependencies needed to run your R package. 

```{r, warning = FALSE, message = FALSE}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
prepare_docker_image()
```
