---
title: "dockr hits CRAN"
output:
  html_document: default
  pdf_document: default
---
```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# dockr

`dockr` 0.8.0 is now available on CRAN. `dockr` is a minimal toolkit to build a 
lightweight Docker image from your R package, in which the package itself is 
installed and available. The image seeks to mirror your R session as close as 
possible with respect to R specific dependencies.

If you want to know, how Docker works, and why you should consider using Docker,
please take a look at the [Docker website](https://www.docker.com/why-docker).

## Installation

Install the development version of `dockr` with:

```{r install_github, eval = FALSE}
remotes::install_github("smaakage85/dockr")
```

Or install the version released on CRAN: 

```{r install_cran, eval = FALSE}
install.packages("dockr")
```

## Workflow

When you work on an R project, it is often desirable to organize the code in an
R package. `dockr` facilitates easy creation of a Docker container image that 
mirrors your R session and includes all of the R dependencies needed to run 
your R package. 

Load the `dockr` package.

```{r load, eval = TRUE}
# load dockr.
library(dockr)
```

```{r setup_cran, echo = FALSE}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```

In order do create the files for the Docker image simply invoke the 
`prepare_docker_image()` function and point to the folder with your package. 
The workflow of `prepare_docker_image()` is summarized below:

1. Build and install the package on your system (just like `devtools::build()`)
2. Identify R package dependencies of the package
3. Detect the version numbers of the loaded and installed versions of these 
packages on your system
4. Write Dockerfile and create all other files for the Docker image

Now, I will let `dockr` do its magic and create the files for a Docker image
container, in which `dockr` is installed together with all of the package 
dependencies, `dockr` needs to run.

```{r, warning = FALSE, message = FALSE}
image <- prepare_docker_image(".")
```

Great, all necessary files for the Docker image have been created, and you 
can build the Docker image right away by following the instructions. It is as 
easy as that! Yeah!

## Files for Docker image

Let us just take a quick look into the folder with the files for the Docker 
image and see the works of `dockr`.

```{r}
list.files(image$paths$dir_image)
```

It contains a Dockerfile and a folder named 'source_packages'.

### Dockerfile

The resulting Dockerfile can be printed with the `print_file()` function:

```{r}
print_file(image$paths$path_Dockerfile)
```

### Folder with Source Packages

The 'source_packages' folder contains the local (non-CRAN) packages, that have 
to be installed in the Docker image container in order for `dockr` to run.

Since `dockr` does not depend on any local (non-CRAN) `source_packages` only 
contains a source package version of `dockr` itself, i.e.:

```{r}
list.files(image$paths$dir_source_packages)
```

## How to edit Dockerfile

If there is need for adding additional lines to the Dockerfile, this can be 
achieved with the `write_lines_to_file()` function. `write_lines_to_file` 
enables you to add new lines to the beginning or the end of the Dockerfile. 

Let us try it out and write a couple of additional lines to the Dockerfile.

```{r}
# write lines to beginning of file.
write_lines_to_file(c("# set maintainer",
                    "MAINTAINER Lars KJELDGAARD <lars_kjeldgaard@hotmail.com>", 
                    ""),
                    image$paths$path_Dockerfile,
                    prepend = TRUE,
                    print_file = FALSE)

# write lines to beginning of file.
write_lines_to_file(c("# check out smaakage85.netlify.com >:-]~~", "smaakage"),
                    image$paths$path_Dockerfile,
                    prepend = FALSE,
                    print_file = TRUE)
```







